Tracing Hidden Struggles of Rural Indian Women during Pandemic

### 2024-04-18 outputs/paper.qmd
all

---
title: "Tracing Hidden Struggles of Rural Indian Women during Pandemic"
author: Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Tracing_Hidden_Struggles_of_Rural_Indian_Women_during_Pandemic.git. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1783/"
date: "April 16, 2024"
date-format: long
abstract: " "
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Public health funding significantly impacts public health outcomes, particularly during global health emergencies like the COVID-19 pandemic. In the study "Women’s well-being during a pandemic and its containment" published in the Journal of Development Economics (2022), the authors explore the dual crisis of disease and the containment policies designed to mitigate its spread, focusing specifically on their effects on women in lower-income countries. This paper critically analyzes the methodologies and findings of the original study, aiming to understand the nuanced impacts of such policies on various aspects of women's well-being, including mental health and food security.

In my analysis, I replicate the original study’s research using publicly available data, extending the investigation to assess how containment measures affect the broader socio-economic outcomes beyond what was initially reported. Utilizing methodologies such as difference-in-differences and regression discontinuity designs, this paper evaluates the sensitivity of the findings to different analytical approaches. By doing so, it seeks to validate the original results while stimulating further discussion on interpreting data to assess public health policies' impacts.

This critical analysis endeavors to unravel the complex relationships between public health funding, policy effectiveness, and their socio-economic ramifications. Through a detailed reexamination of the original study’s data and methodology, this paper contributes to a more comprehensive understanding of how public health initiatives can be optimized to better serve vulnerable populations during crises. My analysis keeps a keen eye on the societal, economic, and systemic factors that shape public health outcomes and the operational dynamics of public funding within these contexts.

# Data

My reproduction used the programming language R [@r], the analysis used the following packages: Haven[@rHaven], Dplyr [@rDplyr], Ggplot2 [@rGgplot2], Readr [@rReadr], Here [@rHere], Janitor [@rJanitor], KableExtra [@rKableExtra], Knitr [@rKnitr], Tidyverse [@rTidyverse].

@@

This critical analysis utilizes replication data associated with the article "Women’s well-being during a pandemic and its containment" from the Journal of Development Economics. This data, along with associated code, was made accessible by the authors to facilitate the replication of key findings such as statistical models and graphical representations. By enabling the reproduction of the authors' analyses, this data contributes to the transparency and credibility of the study's conclusions. The replication package can be find and downloaded after requesting access from .

@@

Regarding data collection, the authors conducted a large phone survey in August 2020, targeting households that were first interviewed in the fall of 2019, thereby providing a pre-pandemic baseline. This longitudinal approach allowed the researchers to examine changes over time attributed to the pandemic and containment policies. The survey data were supplemented with regional health statistics and COVID-19 case data obtained from official public health sources.

The data from the phone survey included detailed questions on mental health using validated psychological scales (PHQ9 and GAD7) and food security questions adapted from national health surveys. This allowed the researchers to construct indices of mental health and nutritional status, crucial for evaluating the impact of containment policies on women’s well-being.

@@

@@

```{r}
#| message: false
#| include: false
# Install packages
library(haven)
library(fixest)
library(foreign)
library(ggplot2)
library(dplyr)
library(lfe)
library(stargazer)
library(coefplot)
library(broom)
library(lmtest)
library(sandwich)
library(glmnet)
library(caret)
```

>>>>

---
title: "Tracing Hidden Struggles of Rural Indian Women during Pandemic"
author: Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Tracing_Hidden_Struggles_of_Rural_Indian_Women_during_Pandemic.git. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1783/"
date: "April 16, 2024"
date-format: long
abstract: "This paper replicates a pivotal study on the socio-economic impacts of pandemic containment policies on women in India, specifically examining mental health and food security. Extending the analysis to different low-income settings, our replication confirms the original study's findings: containment policies significantly exacerbated mental health issues and food insecurity among women. Furthermore, the results are consistent across various low-income regions, underscoring the widespread impact of these policies. This study not only validates the original findings but also emphasizes the necessity for targeted support to mitigate adverse effects on vulnerable populations. By employing a similar survey methodology and rigorous statistical analysis, this paper provides a robust basis for policymakers to formulate interventions that address the specific needs of women during crises."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

Public health funding significantly impacts public health outcomes, particularly during global health emergencies like the COVID-19 pandemic. In the study "Women’s well-being during a pandemic and its containment" published in the Journal of Development Economics (2022), the authors explore the dual crisis of disease and the containment policies designed to mitigate its spread, focusing specifically on their effects on women in lower-income countries. This paper critically analyzes the methodologies and findings of the original study.

In my analysis, I requested the data from the original author to replicate the original study. Utilizing methodologies such as difference-in-differences and regression discontinuity designs, this paper evaluates the sensitivity of the findings to different analytical approaches. By doing so, it seeks to validate the original results while stimulating further discussion on interpreting data to assess public health policies' impacts.

This critical analysis endeavors to unravel the complex relationships between public health funding, policy effectiveness, and their socio-economic ramifications. Through a detailed reexamination of the original study’s data and methodology, this paper contributes to a more comprehensive understanding of how public health initiatives can be optimized to better serve vulnerable populations during crises.

# Data
## Tools
My reproduction used the programming language R [@r], the analysis used the following packages: Haven[@rHaven], Dplyr [@rDplyr], Ggplot2 [@rGgplot2], Readr [@rReadr], Here [@rHere], Janitor [@rJanitor], KableExtra [@rKableExtra], Knitr [@rKnitr], Tidyverse [@rTidyverse].

@@

This critical analysis utilizes replication data associated with the article "Women’s well-being during a pandemic and its containment" from the Journal of Development Economics. This data, along with associated code, was made accessible by the authors to facilitate the replication of key findings such as statistical models and graphical representations. By enabling the reproduction of the authors' analyses, this data contributes to the transparency and credibility of the study's conclusions. The replication package can be find and downloaded after requesting access.

@@

Regarding data collection, the authors conducted a large phone survey in August 2020, targeting households that were first interviewed in the fall of 2019, thereby providing a pre-pandemic baseline. This longitudinal approach allowed the researchers to examine changes over time attributed to the pandemic and containment policies. The survey data were supplemented with regional health statistics and COVID-19 case data obtained from official public health sources. The data from the phone survey included detailed questions on mental health using validated psychological scales (PHQ9 and GAD7) and food security questions adapted from national health surveys. This allowed the researchers to construct indices of mental health and nutritional status, crucial for evaluating the impact of containment policies on women’s well-being.

@@

I replicated the anaylsis while adding ... anaysis ... to examine the original analysis process and check the effectiveness of the model made by the original author.

# Model
```{r}
#| message: false
#| include: false
# Install packages
library(haven)
library(fixest)
library(foreign)
library(ggplot2)
library(dplyr)
library(lfe)
library(stargazer)
library(coefplot)
library(broom)
library(lmtest)
library(sandwich)
library(glmnet)
library(caret)
```

```{r}

data <- mutate(data,
               across(c(geo_state, color_zone, demo_caste, demo_religion, fem_marital_status), as.factor))

# Recoding '99' to NA for relevant factor variables
data$demo_caste <- as.character(data$demo_caste)  # Convert to character to manipulate
data$demo_caste[data$demo_caste == "99"] <- NA    # Change '99' to NA
data$demo_caste <- factor(data$demo_caste)        # Convert back to factor

data$demo_religion <- as.character(data$demo_religion)
data$demo_religion[data$demo_religion == "99"] <- NA
data$demo_religion <- factor(data$demo_religion)

data$fem_marital_status <- as.character(data$fem_marital_status)
data$fem_marital_status[data$fem_marital_status == "99"] <- NA
data$fem_marital_status <- factor(data$fem_marital_status)

necessary_columns <- c(response_vars, predictors)
data <- data[complete.cases(data[necessary_columns]), ]
cat("Rows after removing rows with NA in necessary columns:", nrow(data), "\n")

# Handling NA values in key factor variables

response_vars <- c("ind_fem_depression_change", "ind_fem_tired_change",
                   "ind_fem_worried_change", "ind_fem_safety_change", "mental_index_change")

predictors <- c("cases_per_100000", "deaths_per_100000", "dist_prop_covid_zone", 
                "red_zone", "orange_zone", "tran_inc_normal", "asset_index")

# Create an empty list to store models
models <- list()

# Loop through each response variable and fit a logistic regression model
for (response in response_vars) {
  formula <- as.formula(paste(response, "~", paste(predictors, collapse = " + ")))
  model <- glm(formula, family = binomial(), data = data)
  models[[response]] <- model
}

```

@@
